No need to know: astrophysics-free gravitational-wave cosmology

### 2024-04-09 src/tex/ms.tex
[showyourwork] overleaf sync

In parallel with this work, \citet{magana_inprep} performed a spectral siren analysis on public \ac{LVK} data using histogram bins defined at fixed locations in source-frame mass to flexibly measure the \ac{BBH} mass distribution \citep[as originally presented in][]{mandel_model-independent_2017,ray_non-parametric_2023}.
Both our method and theirs employ \acp{GP}.
In our approach, the black hole mass function is \textit{itself} described by a Gaussian process, with no predefined features.
In \citet{magana_inprep}, on the other hand, the mass distribution is fundamentally a binned piecewise constant function, with a \ac{GP} prior governing the relative heights of these bins.
While this binned model is flexible, sharp bin edges at predefined locations nevertheless constitute unphysical features.
These features may yield overly optimistic constraints on cosmological parameters, possibly explaining the increase in precision that \citet{magana_inprep} find relative to the parametric analysis on the exact same dataset performed in \citet{gwtc3_cosmo}.
\citet{magana_inprep} also include a  binned reconstruction of the \ac{BBH} merger rate as a function of redshift, which we fix or model parametrically in our own work.
We stress, though, that this is not an intrinsic limitation of our method; one could straightforwardly adopt a second Gaussian process to govern the overall scaling of the merger rate with redshift.

% While both the method presented in this work and that of \citet{magana_inprep} employ \acp{GP}, \citet{magana_inprep} uses \acp{GP} to modulate histogram bin heights, leading to a piecewise constant function in log mass, whereas we define the \ac{GP} at each data point, constructing a smooth function with no predefined features in source-frame mass. 
% Sharp bin edges may lead to optimistic constraints on cosmological parameters, especially as such features are unlikely to exist in the true mass distribution, possibly explaining the increase in precision relative to parametric methods seen in \citet{magana_inprep}.

>>>>

In parallel with this work, \citet{magana_hernandez_beyond_2024} performed a spectral siren analysis on public \ac{LVK} data using histogram bins defined at fixed locations in source-frame mass to flexibly measure the \ac{BBH} mass distribution \citep[as originally presented in][]{mandel_model-independent_2017,ray_non-parametric_2023}.
Both our method and theirs employ \acp{GP}.
In our approach, the black hole mass function is \textit{itself} described by a Gaussian process, with no predefined features.
In \citet{magana_hernandez_beyond_2024}, on the other hand, the mass distribution is fundamentally a binned piecewise constant function, with a \ac{GP} prior governing the relative heights of these bins.
While this binned model is flexible, sharp bin edges at predefined locations nevertheless constitute unphysical features.
These features may yield overly optimistic constraints on cosmological parameters, possibly explaining the increase in precision that \citet{magana_hernandez_beyond_2024} find relative to the parametric analysis on the exact same dataset performed in \citet{gwtc3_cosmo}.
\citet{magana_hernandez_beyond_2024} also include a  binned reconstruction of the \ac{BBH} merger rate as a function of redshift, which we fix or model parametrically in our own work.
We stress, though, that this is not an intrinsic limitation of our method; one could straightforwardly adopt a second Gaussian process to govern the overall scaling of the merger rate with redshift.
